---
title: "introVersionControl"
output: html_document
date: "2023-02-07"
---

##STARTING TERMINAL##

bash
cd /Users/slimjims/master/R-studio/Bio197
Ls -F

##Git Access Code##
#Required if you need to push a file: 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

## Version Control ##

Version Control is a powerful tool for keeping track of ALL changes to a document. Especially documents from other users. They are used commonly alongside git and GitHub.
You can use them to keep track of many types of files: text, rscript, python, csv, etc..


## Conflicts ##

Conflicts are created when any user edits the same line in a document. This can be generated by the owner of the document as well.


## What is a commit? ##

This saves the changes made to a file to a repository. It does not have to be all files at a time.

## What is a repository? ##

A respository is a store of all files, and files versions. It includes the history of changes 

## Difference between old and modern version control systems ##

- centralization vs distribution

## Setting up Git

##Create branches
Create with codes such as:
  git branch
  git checkout -    create new branch
  git add      -    track new or old files
  git commit   -    commit to tracking current selection
  
  
  
```{r Intro}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
