---
title: "wrangling-portal"
author: "James Waterford"
date: "2023-03-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Portal, Arizona is a source of a long-term monitoring and experimental ecosystem.



```{r Load-in}
library(readr)
surveys <- read_csv("../197-raw_storage/surveys.csv")
species_data <- read.csv( "../197-raw_storage/species.csv")
plots_data <- read.csv("../197-raw_storage/plots.csv")

head(species_data)
head(surveys)

```

## Combining Data Sets

These data sets are supposed to communicate with each other.
We are going to use the dplyr packages to get to work.

Hindfoot length is measured in *millimeters*, but I want a new variable that shows Hindfoot length in *centimeters*.

`Select` serves the purpose of pulling columns from a data frame.

`mutate` creates a new column, based on the values in another. 
```{r dplyr, echo=FALSE}
library(dplyr)
surveys1 <- select(surveys, year, month, day, species_id)
head(surveys1)

surveys2 <- (surveys, year, species_id, weight)
mutate(surveys,hindfoot_length_cm = hindfoot_length/10)
arrange(surveys,year)
```

Then we do some other things here I guess

```{r}
filter(surveys, species_id != "DS", year > 1995)
filter(surveys, !is.na(weight))
filter(surveys, weight >200)
filter(surveys, weight <=100, sex == "F")
filter(surveys, weight <=100 | sex == "F")

```

```{r}

```

