---
title: "JW_Function_basics"
author: "James Waterford"
date: "2023-03-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functions

THe basic form of a function is to 
`function(a, b) {
c <- a+b
}`

```{r error = TRUE}
name_of_the_function <- function(data_frame, mass, length)
  data_frame %>% 
  group_by(mass, length) %>% 
  summarize(n=n())
```

Exercise 1
-------

```{r Ex1}

convert_pounds_to_grams <- function(pounds) {
  grams = 453.6 * pounds
  return(grams)
}
convert_pounds_to_grams(pounds = 30)


```

Exercise 2: **The weight of dinosaurs**
-----------
The length of an organism is strongly correlated with its body mass. 

```{r dinosaurs}
# Create a function for measuring length from body mass #
length_theropoda <- function(length){
  mass <- 0.73 * length ^ 3.63
  # when length is provided to the function, it will output the mass, calculated by the length #
  return(mass)
}
length_theropoda(length = 32)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
