---
title: "Carribean_Biogeography"
author: "James Waterford"
date: "2023-04-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)

```

##Carribean biogeography as a result of human expansion.


Citation: Gleditsch, Jason et al. (2022), Data for: Contemporizing island biogeography theory with anthropogenic drivers of species richness, Dryad, Dataset, https://doi.org/10.5061/dryad.vx0k6djtn

This data set has been retrieved from the DRYAD public domain, and released by Temple University.

The aim of this database was to measure the *species richness* and *habitat diversity* among different sites across the carribean islands. 

This dataset includes [#] of species, and [#] sites across [#] states.

```{r import}

## I need to move these files ##

a <- "../197-raw_storage/doi_10.5061_dryad.vx0k6djtn__v6/Gleditsch_et_al_Caribbean_Bank_Data.csv"
b <- "../197-raw_storage/doi_10.5061_dryad.vx0k6djtn__v6/Gleditsch_et_al_Caribbean_Herp_SR.csv"
```

```{r csv function}
a_table <- read.csv(a)
b_table <- read.csv(b)

b_table$Bank <- str_to_title(b_table$Bank)
carribean_set <- right_join(a_table,b_table, by = "Bank")
str(carribean_set)
```

```{r}
b_table %>% 
  group_by(Clade) %>% 
  summarise(Count_of_Introduced = sum(Introduced_SR))
b_table %>% 
  group_by(Bank) %>% 
  summarise(Count_of_Introduced = sum(Introduced_SR))
```
```{r}
b_table %>% 
  group_by(Clade) %>% 
  summarise(Count_of_Native = sum(Native_SR))
b_table %>% 
  group_by(Bank) %>% 
  summarise(Count_of_Native = sum(Native_SR))
```


