---
title: "JW_if-else"
author: "James Waterford"
date: "2023-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)

```

## Logical Conditions ##

These are pieces of code that return as `TRUE` or `FALSE` values. 

The common operators include symbols such as: `!=` ,  `==` ,  `>=` ,  `<=`



Practice Logical & Conditional Statements
-------------

We also have some logical functions:

- `is.na()` - tests if a value is `NA`.
- `is.vector` - tests if variable is an array.
- `is.data_frame` - give a guess.
- `which()` - gives the position of all TRUE values

```{r}
letters == "r"
which(letters == "r")
```
Applying Logicals
------------
```{r}

w <- 10.2
w > 10

colors <-  c("red","blue","green")
which(colors == "green")

dna1 <- "attattaggaccaca" 
dna2 <- "attattaggaacaca"
dna1 == dna2
dna2 != dna1


```

## Basic If-Statements ##

If one variable is true, we can create a function based on the the logical value.
```{r}
age_class <- "sapling"
if(age_class == "sapling"){
  y <- 10.0
  y
}
```

## Handling multiple choices ## 

```{r}
age_class <- "seedling"
if(age_class == "sapling"){
  y <- 10.0
} else {
  y <- 5.0
}
  y
 
```
```{r}
age_classing <- function(age_class){
  if(age_class == "sapling"){
   (y <- 10.0)
} else if(age_class == "seedling"){
   (y <- 5.0)
} else if(age_class == "adult"){
    (y <- 20.0)
} else if(age_class == "mature"){
    (y <- 40.0)
} else if(age_class == "established"){
    (y <- 50.0)
}
  y
}
age_classing("established")
```


