---
title: "Open Tree of Life -- 'rotl' "
output: html_document
date: "2022-12-29"
---

```{r setup, include=FALSE}
library(rotl)
```

## Intro

I know nothing about this program, so as I go through, I'll put down codes that I'll be working on. Hopefully this doesn't take too long.

'rotl' helps us interface with Open Tree of Life for Taxonomies and Studies

## Taxa Search

Test the ability of rotl to recognize different taxas.
tnrs_matchnames()

```{r Apes}
apes <- c("Pongo", "Pan", "Gorilla", "Hoolock", "Homo", "Hylobates", "Nomascus")
ape_taxa <- tnrs_match_names(apes)
ape_taxa
```

## Make a Tree

You can also embed plots, for example:
'ott_id' pulls from the ape taxa we made earlier
'plot()' is a necessary component

```{r tree, echo=FALSE}
tr <- tol_induced_subtree(ott_ids = ott_id(ape_taxa))
plot(tr)
```

## What if 'trns_match_names' assigns incorrect values? 




``` {r idk}
inspect(ape_taxa, taxon_name = 'Hylobate')
update(ape_taxa, taxon_name = "Hylobate",
       new_ott_id = 166552)

```


